---
title: "52414: Lab 2"
author: "Gil Shiloh and Elky Sandor"
date: "June 1, 2021"
output: html_document
---


## *Lab 2: Visualization Through `ggplot`*  
<br/><br/>  
  

  
  



#### Background: 

You've been hired as a data analyst at at football (soccer) club. 
Since this is a small and under-funded club, you will not have access to real-football data, but to data from 
the football computer game fifa18. Your job is to analyze this dataset and extract meaningful insights from the data in order 
to help your club make better decisions. 

#### Data File: 
You will load and analyze the fifa18 football dataset file called "fifa_data.csv". <br> 
The dataset contains detailed information about each player in the game, including: names, age, nationality, overall ability, estimated potential ability, current club and league, market value, salary (wage), ability at different football skills (also called 'attributes', e.g. Ball.control, Sprint.speed ...), ability to play at different position in the game (CF, CM, ...) and the preferred positions of the player. 



Required Libraries:
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(dplyr)
library(corrplot)
library(scales)   # needed for formatting y-axis labels to non-scientific type
library(radarchart)
library(tidyr)
library(tidyverse)
library(reshape2) # melt
library(ggthemes)
library(rworldmap) # world map
library(modelr)
library(radarchart) #Spider chart
############################################
library(e1071) #Q1.c -  skewness() and kurtosis()
library(grid) # geom_segment
library(ggrepel)# Use ggrepel::geom_label_repel
library(fmsb)  #Spider chart

options("scipen"=100, "digits"=4)  # avoid scientific display of digits. Take 4 digits. 

```




```{r}
# A measure of category's diversity
DIV <- function(category_vec){
  t <- table(category_vec)
  p <- t/sum(t)
  return(sum(p^2))
}

cleaned_data <- fifa_players %>% select(Nationality,Club) %>% na.omit()

number_of_nationality_in_club <- cleaned_data %>% group_by(Club, Nationality) %>% summarise(count = n()) %>% group_by(Club) %>% summarise(N_nation=n()) %>% arrange(desc(N_nation)) %>% mutate(Club = factor(Club, level=unique(Club)))

DIV_in_club <- cleaned_data %>% group_by(Club) %>% summarise(DIV = DIV(Nationality))%>% arrange(DIV)%>% mutate(Club = factor(Club,level=unique(Club)))  # arrange(desc(DIV)) %>% 

# Plot number of different nationalities in each club
g <- ggplot(data = number_of_nationality_in_club %>% head(8), aes(x = Club, y = N_nation,color = Club))
g + geom_bar(stat="identity")

# Plot DIV (diversity?) of different nationalities in each club
g <- ggplot(data = DIV_in_club %>% head(12),aes(x = Club,y = DIV, color = Club))
g <- g + geom_bar(stat="identity")
g
```


PLEASE ADD YOUR SOLUTION BELOW, WITH A CLEAR SEPARATION BETWEEN THE PARTS!


## Q1 basic statistic

#### a

*Make a plot showing the overall ability distribution of all players,How skewed is the distributions? does it have fat tails? Plot on top of the overall distribution a Normal distribution matching its first two moments.*

```{r, warning=FALSE}
ggplot(data = fifa_players,aes(x=Overall))+
  geom_density(alpha= .3,fill="#00BFC4")+
  stat_function(fun = dnorm,args = list(mean = mean(fifa_players$Overall),sd =sd(fifa_players$Overall)),col="orange",lwd=0.5 )+
  scale_x_continuous(name = "players overall",breaks = breaks_width(5))
skewed_Wage <- skewness(fifa_players$Overall)
skewed_Wage
fat_tail_ind<-kurtosis(fifa_players$Overall)
fat_tail_ind
```

*The results are that the distribution is 0.008293 skewed and have kurtosis of -0.02102*
*from the low skewed we can tell that the distribution of the overall is pretty symmetry.*
*and from the low forth moment (compared to normal distribution) we can tell that the tails are thin*
*we can see that the normal distribution above the Overall distribution is almost the same which means that the normal distribution described good the data*


#### b

*Make a plot comparing the multiple overall ability distributions of players according to the continent of the players. Describe which continents have especially good/bad players.*

```{r, warning=FALSE}
ggplot(data = fifa_players,aes(x=Overall,y=Continent,color = Continent ))+
  geom_boxplot()+
  labs(title = "Players overall quality by continent")
```

*From the graph we can see that Antartica have extremely bad soccer players, on the other side Africa and South America have great players*

#### c

*Make a plot showing the density of players’ value distribution.*

```{r, warning=FALSE}
ggplot(data = fifa_players,aes(x=Value))+
  geom_density(alpha= .1,fill="#00BFC4")+
  scale_x_continuous(name = "players value")
```

*Next, make a separate plot showing the density distribution of the log of players’ value.*

```{r, warning=FALSE}
ggplot(data = fifa_players,aes(x=Value))+
  geom_density(alpha= .2,fill="#00BFC4")+
  scale_x_continuous(trans = "log")
```

*The second graph (log of value) is a better visualization because in the first graph the values are moving too quickly to understand them in a small graph, on the other hand in the second (log) graph, the scale is the log of the the original values, which reduces very much the differences of the original values. so the data grow in a clearly rate which is easier to see and understand*

#### d

*Are the top-10 players with the highest value also the top-10 best players in terms of overall ability? Show tables for both and compare. Who is the best player not in the top-10 valued players?*

```{r, warning=FALSE}
top_10_val <- fifa_players %>% arrange(desc(Value))%>%
  dplyr::select(Name,Overall, Value)%>%
  head(10)
top_10_val
```

```{r, warning=FALSE}
top_10_over <- fifa_players %>% arrange(desc(Overall))%>%
  dplyr::select(Name,Overall,Value)%>%
  head(10)
top_10_over
```
```{r, warning=FALSE}
compare<-anti_join(top_10_over,top_10_val,by = c("Name", "Overall", "Value"))
compare[which.max(compare$Overall),]
```

*We can see that Neuer is the best player not in the top-10 valued players because in soccer the GK is a position that normally is under valued compare to all the others*

#### e

*Show a table of the 10 youngest and 10 oldest teams in terms of average players age*

```{r, warning=FALSE}
ave_age_by_team<-aggregate(Age~Club,data = fifa_players,FUN = mean)
youngest_teams<-head(arrange(ave_age_by_team,Age),10)
knitr::kable(youngest_teams, caption = "youngest teams")

```

```{r, warning=FALSE}
oldest_teams<-head(arrange(ave_age_by_team,desc(Age)),10)
knitr::kable(oldest_teams, caption = "oldest teams")

```

## Q2

#### a

*Plot the average potential ability by age of all players, for players 35 years old or younger*

```{r, warning=FALSE}
under_35<-filter(fifa_players,Age<=35)
potential_by_age<-aggregate(Potential~Age,data = under_35,FUN = mean)
ggplot(potential_by_age,aes(x=Age,y=Potential))+
geom_line(color = "red")+
geom_point()+
scale_x_continuous(name = "Age",breaks = breaks_width(5))+
labs(title = "average potential by age")

```

#### b

*Plot the average difference between a player’s overall ability to potential ability as a function of age, up to age 35*

```{r, warning=FALSE}
Diff<-abs(under_35$Overall-under_35$Potential)
under_35<-mutate(under_35,Diff)
aver_diff<-aggregate(Diff~Age,data = under_35,FUN = mean)
ggplot(aver_diff,aes(x=Age,y=Diff))+
  ylab("difference between overall and potential")+
  geom_line(color="red")+
  geom_point()
```

*From the graph we can understand that as young as the player it's better for future development until the age of 25 because after that the Difference between the Overall and the Potential becomes under 2.5 so the player overall become more stabilized.*

#### c

*We are seeking young (age≤21) players with high potential (>70). Show a scatter plot of these players comparing their potential ability (x-axis) and current value (y-axis).Find the 10 most-undervalued players, i.e. having the lowest value compared to their predicted value by potential using a simple linear regression model. Calculate for each of them what is a fair value matching their potential that you be willing to pay in order to by them to your club and show these 10 players with their name, age, overall ability, potential, actual value and fair value it a table.*

```{r, warning=FALSE}
good_young<-filter(under_35,Age<=21&Potential>70)
regg_1<-lm(Value~Potential,data = good_young)
predicted_val<-regg_1$fitted.values
ggplot(good_young,aes(x=Potential,y=Value))+
  geom_point(color="blue")+
  geom_line(aes(y=predicted_val))
```
```{r}
preper_under_val<-predicted_val-good_young$Value
index<-which(preper_under_val%in%head(sort(preper_under_val,decreasing = TRUE),10))
under_val_players<-as.vector(good_young$Name[index])
fair_val<-predicted_val[index]
under_val<-good_young$Value[index]
small_under_val<-good_young[index,]
small_under_val<-mutate(small_under_val,fair_val)%>%
  dplyr::select(Name,Age,Overall,Potential,Value,fair_val)
knitr::kable(small_under_val, caption = "top 10 under valued player")
```

*As we can see in the question above we need to find the 10 most-undervalued players. which means that we need to take the players with the max difference between their value and their predict value, but only when the predict value is higher then the value(taking the other option will be overvalue players). So as we can see those are the players that we show in our table.*

#### d

*Use the rworldmap package to display the world map and color each country based on the median potential of players from this nationality.*

```{r, warning=FALSE}
prep_to_map<-aggregate(Potential~Nationality,data = fifa_players,FUN = median)
mapped_data <- joinCountryData2Map(prep_to_map, joinCode = "NAME", nameJoinColumn = "Nationality")
theMap <- mapCountryData(mapped_data, nameColumnToPlot="Potential",catMethod ="fixedWidth",colourPalette = "palette",
oceanCol = "lightblue", missingCountryCol = "white",mapTitle = " median potential of players per country"
,aspect = "variable")
```

#### e

*Repeat the above analysis but this time display a world map where each country is colored by the median ratio of potential to value of players. Find an under-valued country you’d recommend to travel to (i.e. a country with cheap players compared to their potential average quality).*

```{r, warning=FALSE}
ratio<-fifa_players$Value/fifa_players$Potential
fifa_map<-mutate(fifa_players,ratio)
prep_to_map_2<-aggregate(ratio~Nationality,data = fifa_map,FUN = median)
mapped_data_2 <- joinCountryData2Map(prep_to_map_2, joinCode = "NAME", nameJoinColumn = "Nationality")
theMap_2 <- mapCountryData(mapped_data_2, nameColumnToPlot="ratio",catMethod ="fixedWidth",colourPalette = "rainbow",
oceanCol = "lightblue", missingCountryCol = "white",mapTitle = " median ratio of players per country"
,aspect = "variable")
```


```{r}
check_1<-aggregate(Potential~Nationality,data = fifa_map,FUN = mean)
check_2<-aggregate(Value~Nationality,data = fifa_map,FUN = mean)
check_3<- full_join(check_2,check_1,by = "Nationality")
check_4<-check_3$Value/check_3$Potential
best_country<-aggregate(ratio~Nationality,data = fifa_map,FUN = mean)
cheap_conti<-top_n(best_country,1,ratio)
knitr::kable(cheap_conti, caption = "recommend country")

```

*We would recommend to travel to Gabon*

## Q3

#### a

*We are interested in finding out which positions are similar in terms of players’ performance. Extract the 26 non-goalkeeper positions (CAM, CB, ..., ST). Calculate the correlation between players’ ability in each pair of positions and show a heatmap correlation-plot of the correlations’ matrix. What three positions have the least average correlations with other skills?*

```{r, warning=FALSE}
no_gk<-fifa_players_positions[!(fifa_players_positions$Preferred.Positions=="GK "),]
cor_positions<-cor(no_gk[,-c(1,28:30)])
corrplot(cor_positions,order='hclust', tl.cex = 0.7,mar = c(0,0,1,0), title="a. Pairwise correlations between positions")
mean_of_cor<-rowMeans(cor_positions)
head(sort(mean_of_cor),3)
```

*Three positions that have the least average correlations with other skills are CB, LCB and RCB.*

*We are interested in finding out which skills are similar in terms of players’ performance at the position. Extract the 29 skills for non-goalkeeper players (Acceleration, …, Volleys, except ’GK.’ skills). Calculate the correlation between players’ ability in each pair of skills and show a heatmap correlation-plot of the correlations’ matrix. What two skills seem least correlated with other skills?*

```{r, warning=FALSE}
ready_to_cor_2<-cor(fifa_players_attribures[,c(2:12,18:35)])
corrplot(ready_to_cor_2,order='hclust', tl.cex = 0.7,mar = c(0,0,1,0), title="a. Pairwise correlations between skills")
mean_of_cor_2<-rowMeans(ready_to_cor_2)
head(sort(mean_of_cor_2),2)
```

*Two skills that are least correlated with other skills are Strength and Jumping.*

#### b

*Consider the following indicators of players performance: overall players’ performance, their potential, their salary (wage) and their market value. Show a correlation-plot of players’ 34 skill levels (Acceleration, …, Volleys) vs. these four indicators. Find the 10 skills with the highest average correlation with the four inidcators and list them in a table.*

```{r, warning=FALSE}
four_indicators<-fifa_players[,c(6,7,10,11)]
skills_dat<-fifa_players_attribures[,c(2:35)]
ready_to_cor_3<-cor(four_indicators,skills_dat)
corrplot(ready_to_cor_3,method= "square",cl.lim = c(-1,1),cl.cex = 0.5,tl.cex = 0.7,mar = c(0,0,1,0), title="a. Pairwise correlations between skills",is.corr = FALSE)
mean_of_cor_3<-colMeans(ready_to_cor_3)
best_skills<-head(sort(mean_of_cor_3,decreasing = TRUE),10)
best_skills<-as.data.frame(best_skills)
knitr::kable(best_skills, caption = "best skills")
```

#### c

*Build a team of 11 different players - For each of the 26 non-goalkeaper positions (26 from above plus goalkeaper, GK), find the player with the best performance at this position.Find the goal keaper (Preffered.Positions is GK) with the best overall performance.From the players obtained above, find 11 distinct players maximizing the average overall performance of the team, with the constraint that there must be a goalkeaper (preferred position GK).List the players in a table including their overall performance and the team average overall score.*
*Next, peak six different players of your choice from your team, one of which is the goalkeaper. Using the function radarchart::chartJSRadar, graph their abilities (individually for all 6 players) in the top 10 skills according to 3.b in a radar chart*

```{r, warning=FALSE}
all_positions <-fifa_players[,c(46:71)]
all_max<-apply(all_positions,2,function(x)  which( x == max(x) ))
index_2<-unlist(all_max)
index_2<-unique(index_2)
best_in_pos<-fifa_players[index_2,]
fifa_players$Preferred.Positions<-as.character(fifa_players$Preferred.Positions)
only_gk<-filter(fifa_players,Preferred.Positions=="GK ")
best_gk<-only_gk[which.max(only_gk$Overall),]
our_team<-rbind(best_in_pos[-which.min(best_in_pos$Overall),],best_gk)
average<-data_frame(Name= "Team average",Overall = mean(our_team$Overall))
our_team<-dplyr::select(our_team,Name,Overall)
final_team<-rbind(our_team,average)
knitr::kable(final_team, caption = "best team")
```



```{r,warning=FALSE}
chosen_players<-rbind(head(arrange(best_in_pos,desc(Overall)),5),best_gk)
data_to_spider<-dplyr::select(chosen_players,rownames(best_skills))
data_to_spider<- rbind(rep(100,dim(data_to_spider)[2]),rep(0,dim(data_to_spider)[2]),data_to_spider)
title<-as.character(chosen_players$Name)
par(mar=rep(1,4))
par(mfrow=c(2,3))
for(i in 1:6){
  radarchart(data_to_spider[c(1,2,i+2),],axistype=1, pcol=rgb(0.8,0.2,0.5,0.9),pfcol = rgb(0.8,0.2,0.5,0.4), plwd=4,plty=1, cgcol="grey",cglty=1, axislabcol="grey", caxislabels=seq(0,20,5), cglwd=0.8,vlcex=0.8,title=title[i])
}
```

#### d

*Repeat the analysis of question 2.a., but this time show the 34 different skills*

```{r, warning=FALSE}
no_gk_under_35<-under_35[!(under_35$Preferred.Positions=="GK "),]
check_pos<-no_gk_under_35[,c(3,46:71)]
mean_check<-aggregate(.~Age,data = check_pos,FUN = mean)
trans_check<-t(mean_check)
vec_to_insert<-c(colnames(under_35[,c(46:71)]))
trans_check<-as.data.frame(trans_check)
colnames(trans_check) <- trans_check[1,]
trans_check<-trans_check[-1,]
trans_check<-mutate(trans_check, position = vec_to_insert)
skills_to_P<-melt(trans_check,id.vars = "position")
colnames(skills_to_P)[2]<-"Age"
skills_to_P$Age<-as.numeric(as.character(skills_to_P$Age))
ggplot(data = skills_to_P,aes(x=Age,y=value,group=position))+
geom_line(aes(color=position),show.legend = FALSE)+
theme(axis.text.x = element_text(angle = 60, hjust = 1))+
facet_wrap(~position)+
 xlim(15,35)+
  labs(title = "Position rate by Age")
```

*As we can see there's no one position who peak at the youngest age, it can be explained by the lack of experience*
*We noticed that the defense players peak at the age of 30, the center players peak at the age of 35 and the attacking players peak at the age of 25.*

#### e

*Fit a multiple regression model predicting player’s overall performance based on their wage and age. Find the 10 players with the highest difference between their overall performance level and the regression model prediction, and list them in a table.*

```{r, warning=FALSE}
regg_2<-lm(Overall~Wage+Age,data = fifa_players)
differnce_2<-regg_2$residuals
index_3<-which(differnce_2%in%head(sort(differnce_2,decreasing = TRUE),10))
top_10_diff<-fifa_players[index_3,]%>% dplyr::select(Name)
top_10_diff<-mutate(top_10_diff,Differnce = differnce_2[index_3])%>%
  arrange(desc(Differnce))
knitr::kable(top_10_diff, caption = "players with highest difference")
```

# Q4

```{r}
# A measure of category's diversity
DIV <- function(category_vec){
  t <- table(category_vec)
  p <- t/sum(t)
  return(sum(p^2))
}

cleaned_data <- fifa_players %>% dplyr::select(Nationality,Club) %>% na.omit()

number_of_nationality_in_club <- cleaned_data %>% group_by(Club, Nationality) %>% summarise(count = n()) %>% group_by(Club) %>% summarise(N_nation=n()) %>% arrange(desc(N_nation)) %>% mutate(Club = factor(Club, level=unique(Club)))

DIV_in_club <- cleaned_data %>% group_by(Club) %>% summarise(DIV = DIV(Nationality))%>% arrange(DIV)%>% mutate(Club = factor(Club,level=unique(Club)))  # arrange(desc(DIV)) %>% 

# Plot number of different nationalities in each club
g <- ggplot(data = number_of_nationality_in_club %>% head(8), aes(x = Club, y = N_nation,color = Club))
g + geom_bar(stat="identity")

# Plot DIV (diversity?) of different nationalities in each club
g <- ggplot(data = DIV_in_club %>% head(12),aes(x = Club,y = DIV, color = Club))
g <- g + geom_bar(stat="identity")
g
```

#### a

*The goal of the both plots is to show which team is the most diverse in terms of players nationality. The first plot represent the highest 8 clubs with players from different number of nations. And The second plot shows the Worst 12 clubs proportion. This measure takes into account the proportions of players from each nation, not just the overall number of nation.*

#### b

The problems that we found:

*1-The first column is NA and those values should be removed.*
*2-There's no title.*
*3-The names of the clubs is stepping on the others names so we can't read them.*
*4-The filling of each column is with the same color, and it makes it hard to understand.*
*5-The names of the axes can be defined better*

#### c

```{r, warning=FALSE}
number_of_nationality_in_club <- cleaned_data %>% group_by(Club,Nationality) %>% summarise(count = n()) %>% group_by(Club) %>% summarise(N_nation=n()) %>% filter(Club!="") %>% arrange(desc(N_nation)) %>% mutate(Club = factor(Club,level=unique(Club)))
```

```{r}
DIV_in_club <- cleaned_data %>% group_by(Club) %>% summarise(DIV = DIV(Nationality))%>%  filter(Club!="") %>% arrange(DIV)%>% mutate(Club = factor(Club,level=unique(Club)))
g <- ggplot(data = number_of_nationality_in_club %>% head(8),aes(x = Club,y = N_nation,color = Club, fill = Club))
g + geom_bar(stat="identity") + 
theme(axis.text.x = element_text(face = "bold", size = 10, angle = 60, hjust = 1)) + ylab('Number of Nations') +
labs(title="top 8 clubs by number of Nationalities")
```

```{r}
g <- ggplot(data = DIV_in_club %>% head(12),aes(x = Club, y = DIV, color = Club, fill = Club))
g + geom_bar(stat="identity") + theme(axis.text.x = element_text(face = "bold", size = 10, angle = 60, hjust = 1)) + ylab('porportion of Nationality') +
labs(title="least 12 clubs in term of porportion of Nationality")
```